Word-Referent Identification Under Multimodal Uncertainty

### 2018-04-13 Writeup/ms.tex
changes on ms

\documentclass[english,man]{apa6}

@@

  \usepackage{ifthen} % Only add declarations when endfloat package is loaded
  \ifthenelse{\equal{\string man}{\string man}}{%
   \DeclareDelayedFloatFlavor{ThreePartTable}{table} % Make endfloat play with longtable
   % \DeclareDelayedFloatFlavor{ltable}{table} % Make endfloat play with lscape
   \DeclareDelayedFloatFlavor{lltable}{table} % Make endfloat play with lscape & longtable
  }{}%

@@

How does the mind deal with uncertainty to identify the speaker's
intended word? To address this question, many researchers in recent
years have adopted the rational approach to cognition (Anderson, 1990),
which corresponds to Marr's computational level of analysis. This
approach has had a tremendous impact not only on speech related research
(Clayards, Tanenhaus, Aslin, \& Jacobs, 2008; Feldman, Griffiths, \&
Morgan, 2009; Kleinschmidt \& Jaeger, 2015; Norris \& McQueen, 2008),
but also on many other disciplines in the cognitive sciences (for
reviews, see Chater \& Manning, 2006; Knill \& Pouget, 2004; Tenenbaum,
Kemp, Griffiths, \& Goodman, 2011). Within this framework, researchers
typically consider an \enquote{ideal observer} model which provides a
characterization of the task/goal and shows what the optimal performance
should be under this characterization. When there is uncertainty in the
input, the ideal observer performs an optimal probabilistic inference.
For example, in order to recognize an ambiguous linguistic input, the
model uses all available probabilistic knowledge in order to maximize
the accuracy of this recognition. Thus, when the task is well specified,
the ideal observer model can be seen as a theoretical upper limit on
performance. It is often used, not so much as a realistic model of human
performance, as much as a baseline against which human performance can
be compared (Geisler, 2003). When there is a deviation from the ideal,
this can reveal extra constraints on human cognition, such as
limitations on working memory or attentional resources.

In this line of research, Clayards et al. (2008) simulated auditory
uncertainty by manipulating the probability distribution of a cue (VOT)
that differentiated similar words (e.g., \enquote{beach} and
\enquote{peach}). They found that humans were sensitive to these
probabilistic cues and their judgments closely reflected the optimal
predictions. In another work, Feldman et al. (2009) studied the
perceptual magnet effect, which is related to how categories influence
the way the stimuli are perceived. They showed that this effect can be
explained as the consequence of optimally solving the problem of
perception under uncertainty. Both studies explored optimal performance
under uncertainty from one modality, i.e., the auditory modality. There
is however extensive evidence that information from the visual modality,
in particular the speaker's facial features, influences the perception
of auditory information (see Campbell, 2008 for a review). Bejjanki,
Clayards, Knill, and Aslin (2011) offered a mathematical
characterization of how cues from speech and lip movements can be
optimally combined and showed that human performance during audio-visual
phonemic labeling was consistent (at least at the qualitative level)
with the behavior of the ideal observer.

The speaker's facial features are not the only visual information that
can be used to disambiguate the auditory input, however. For instance,
in a referential context, information about the identity of the semantic
referent can be integrated with linguistic information to resolve
lexical and syntactic ambiguities in speech (e.g., Eberhard,
Spivey-Knowlton, Sedivy, \& Tanenhaus, 1995; Spivey, Tanenhaus,
Eberhard, \& Sedivy, 2002; Tanenhaus, Spivey-Knowlton, Eberhard, \&
Sedivy, 1995). To our knowledge, however, no study offered an ideal
observer analysis of word identification in such context, that is, when
the mind has to combine cues from the sound and the referent. Imagine,
for example, that someone is uncertain whether they heard \enquote{pea}
or \enquote{bee}, does this uncertainty make them rely more on the
referent (e.g., the object being pointed at)? Vice versa, if they are
not sure if they saw a bee or a fly, does it make them rely more on the
sound? More importantly, when input in both modalities is uncertain to
varying degrees, do they weight each modality according to its relative
reliability (which is the optimal strategy), or do they over-rely on a
particular modality (which is a sub-optimal strategy)?

The visual information provided by the referent is arguably different
from the visual information provided by the speaker's facial features
(such as lip movements). There seems to be at least two fundamental
differences between these two cases, and both can influence the cue
combination strategy:

First, in the case of audio-visual speech, modalities are understood to
differ only in terms of their reliability. In a referential context,
however, modalities also differ in terms of their roles in the
referential process: the auditory input represents the symbol (i.e., the
word) whereas the visual input represents the referent. It has been
suggested that because of its referential property, speech is a
privileged signal for humans, starting in infancy (see Vouloumanos \&
Waxman, 2014 for a
review).\footnote{There is a debate over whether speech is privileged for infant and adults for the same reasons. Whereas Waxman and colleagues suggest that speech is privileged for both infants and adults because of its ability to refer, Sloutsky and colleagues suggest that speech might not have a referential status from the start. Rather, according to Sloutsky and colleagues, speech seems to be prefered by infants only because of a low level auditory "overshadowing".}
Thus, in a referential context, it is possible that we do not treat the
auditory and visual modalities as equivalent sources of information.
Instead, there could be a bias for the auditory modality beyond what is
expected from informational reliability alone.

Second, in the case of audio-visual speech, the associations are
expected to be perceptually correlated. The expectation for this
correlation is such that when there is a mismatch between auditory and
visual input, people still integrate them into a unified (but illusory)
percept (McGurk \& MacDonald, 1976). In the case of referential
language, however, the multimodal association is by nature
\emph{arbitrary} (Greenberg, 1957; Saussure, 1916). For instance, there
is no logical/perceptual connection between the sound \enquote{bee} and
the corresponding insect. Moreover, variation in the way the sound
\enquote{bee} is pronounced is generally not expected to correlate
perceptually with variation in the shape (or any other visual property)
in the category of bees. Thus, one could imagine that cue combination in
the case of arbitrary audio-visual associations (word-referent) is less
automatic, more effortful, and therefore less conducive to optimal
integration than it is in the case of perceptually correlated
associations (as in audio-visual speech perception).

In the current study, we investigate how people combine cues from the
auditory and the visual modality to recognize words in a referential
context. In particular, we study how this combination is performed under
various degrees of uncertainty in both the auditory and the visual
modality. We perform a rational analysis of the task. First we propose
an ideal observer model that performs the combination in an optimal
fashion. Second we compare the predictions of the optimal model to human
responses. Humans can deviate from the ideal for several reasons. For
instance, as mentioned above, a sub-optimality can be induced by the
suggested privileged status of speech or by the arbitrariness of the
referential association. In order to study possible patterns of
sub-optimality, we compare the optimal model (which provides a normative
benchmark) to a descriptive model which is identical to the optimal
model, except that the parameters in the former are fitted to human
responses.

@@

\section{Models and predictions}\label{models-and-predictions}

In this section we describe the multimodal cue combination model. First,
we briefly introduce the experimental paradigm, and second we explain
how behavior in this paradigm can be characterized optimally with an
ideal observer model.

@@

We introduce a new task that tests word recognition in a referential
context. We use two visual categories (cat and dog) and two auditory
categories (/b/ and /d/ embedded in the minimal pair /aba/-/ada/). For
each participant, an arbitrary pairing is set between the auditory and
the visual categories, leading to two audio-visual word categories
(e.g., dog-/aba/, cat-/ada/). In each trial, participants are presented
with an audio-visual target (the prototype of the target category),
immediately followed by an audio-visual test stimulus
(Figure~\ref{fig:task}). The test stimulus may differ from the target in
both the auditory and the visual components. After these two
presentations, participants press \enquote{same} or \enquote{different}.

@@

This task is similar to the task introduced by Sloutsky and Napolitano
(2003), and used in subsequent studies in both children and adults (see
Robinson \& Sloutsky, 2010 for a review). However, in this original
task, participants are asked whether or not the two audio-visual
presentations are \emph{identical}. Such task would allow us to probe
audio-visual encoding, but not necessarily language understanding. The
latter requires---in addition to the perceptual encoding---a categorical
judgement, i.e, determining whether or not two similar tokens are
members of the same phonological/semantic category. Thus, our task is
category-based: Participants are asked to press \enquote{same} if they
think the second item (the test) belonged to the same category as the
first (target) (e.g., dog-/aba/), even if there is a slight difference
in the sound, in the referent, or in both. They are instructed to press
\enquote{different} only if they think that the second stimulus was an
instance of the other word category (cat-/ada/). The task also includes
trials where pictures are hidden (audio-only) or where sounds are muted
(visual-only). These unimodal trials provide us with the participants'
evaluation of the probabilistic information present in the auditory and
visual categories. This unimodal distributions are used as inputs to the
optimal combination model, described below.

@@

The ideal observer model combines probabilistic information from the
auditory and visual modalities, and makes a categorical decision. How
can this task be characterized in a formal way, and what would be the
optimal behavior? If the stimuli were continuous (e.g., size, location),
cues derived from each modality would be characterized with their
sensory noise variance only, and (assuming conditional independence)
their combined information would be characterized with a simple linear
combination rule where the combined estimate is an average of the
individual estimates, each weighted according to its relative
reliability (Alais \& Burr, 2004; Ernst \& Banks, 2002). However, in our
case decision is made over categorical variables (phonemes and
concepts), the optimal model should take into account, not only the
noise variability around an individual perceptual estimate, but also its
categorical variability (i.e., the probability with which the estimate
belongs to a given category).

One way to account for the categorical variability is to model the
decision step which maps sensory input onto task-dependent categories
(Bankieris, Bejjanki, \& Aslin, 2017; Bejjanki et al., 2011). We
represent the decision step as follows (see Figure~\ref{fig:model} for
illustration). If we assume that both the visual and the auditory cues
vary along a single dimension, an audio-visual signal \((a,v)\) can be
represented as a point in this two-dimensional space, and a word
category can be characterized with a bivariate probability distribution
over audio-visual
tokens.\footnote{For the sake of clarity, we assume here that the sensory noise around this point is small/negligible compared to the categorical variability. But see below how the noise can be integrated when its effect is relatively large}
Each category is peaked around the most typical audio-visual token, and
the probability of membership in the category (the likelihood
probability) decreases gradually as the auditory token deviates from the
typical sound instance (depending on the variance of the auditory
category) and/or as the visual token deviates from the typical visual
instance (depending on the variance of the visual category).

A crucial feature of the ideal observer model is that it makes use of
the entire probability distribution of a given category in order to make
optimal decisions (e.g., Clayards et al., 2008). This means that the
probability of picking a word (the posterior probability) should mirror
closely the probability of the combined information (the likelihood
probability). For instance, the model becomes less ambiguous and more
certain as the tokens get closer to the typical instance of the
category. In what follows, we formalize this sense of optimality using a
Bayesian framework.

An auditory category \(A\) and a visual category \(V\) are defined as
distributions in the auditory and the visual dimensions, respectively. A
word category \(W\) is defined as a bivariate distribution in the
audio-visual. More specifically, we define it as the joint distribution
of the auditory and the visual categories. For simplicity, we assume
both auditory and visual categories to be normally distributed. That is,
if \(w=(a,v)\) is an audio-visual token, then:
\[ p(a | A) \sim  N(\mu_A, \sigma^2_A) \] where \(\mu_A\) and
\(\sigma^2_A\) are respectively the mean and the variance of the
auditory category, \[ p(v | V) \sim  N(\mu_V, \sigma^2_V) \] where
\(\mu_V\) and \(\sigma^2_V\) are the mean and the variance of the visual
category, and

\[ p(w | W) \sim  N(M_W, \Sigma_W) \] where \(M_W=(\mu_A, \mu_V)\) and
\(\Sigma_W\) are mean and the covariance matrix of the word category. We
assume that auditory and visual variables are independent (i.e.,
uncorrelated), so the covariance matrix is simply: \[
   \Sigma_W=
  \left[ {\begin{array}{cc}
   \sigma^2_A & 0 \\
   0 & \sigma^2_V \\
  \end{array} } \right]
\]

{\centering \includegraphics[width=400px]{pictures/model}

\caption{Illustration of the model using simulated data. A word category is defined as the joint bivariate distribution of an auditory category (horizontal, bottom panel) and a visual semantic category (vertical, left panel). Upon the presentation of a word token $w$, participants guess whether it is sampled from the word type $W_1$ or from the word type $W_2$. Decision threshold is where the guessing probability is 0.5.}\label{fig:model}
\end{figure}

This assumption simply says that, given a word-object mapping, e.g.,
W=(\enquote{cat}-CAT), variation in the way \enquote{cat} is pronounced
does not correlate with changes in any visual property of the object
CAT, which is a valid
assumption.\footnote{Note that this assumptions is more adequate in the case of arbitrary associations such as ours, and less so in the case of redundant association such as audio-visual speech. In the latter, variation in the pronunciation is expected to correlate, at least to some extent, with lip movements.}

Now we turn to the crucial question of modeling how optimal decision
should proceed based on the probabilistic information in the auditory
and the visual modalities. In our task, we have two word categories:
dog-/aba/ (\(W_1\)) and cat-/ada/
(\(W_2\)).\footnote{This mapping is randomized in the experiments.}
Participants can be understood as choosing one of these two word
categories (Figure~\ref{fig:model}). For an ideal observer, the
probability of choosing category 2 when presented with an audio-visual
instance \(w=(a,v)\) is the posterior probability of this category: \[
p(W_2 | w)=\frac{p(w|W_2)p(W_2)}{p(w|W_2)p(W_2)+p(w|W_1)p(W_1)}
\] Using our assumption that the cues are uncorrelated, we have
\[p(w | W) = p(a,v| W) = p(a| A)p(v| V)\] Under this assumption, the
posterior probability reduces
to:\footnote{See the appendix for the details of the derivation}

\begin{equation}
 p(W_2 | w)=\frac{1}{1+(1+\epsilon)\exp(\beta_0+\beta_aa+\beta_vv)}
\end{equation}

where \[1+\epsilon=\frac{p(W_1)}{p(W_2)}\]
\[\beta_0=\frac{\mu^2_{A2}-\mu^2_{A1}}{2\sigma^2_{A}}+\frac{\mu^2_{V2}-\mu^2_{V1}}{2\sigma^2_{V}}\]

\[\beta_a=\frac{\mu_{A1}-\mu_{A2}}{\sigma^2_{A}}\]
\[\beta_v=\frac{\mu_{V1}-\mu_{V2}}{\sigma^2_{V}}\]

The parameter \(\epsilon\) represents the differential between the
categories' prior probabilities. However, since the identity of word
categories is randomized across participants, \(\epsilon\) measures,
rather, a response bias to \enquote{same} if \(\epsilon > 0\), and a
response bias to \enquote{different} if \(\epsilon < 0\). We expect a
general bias towards answering \enquote{different} because of the
categorical nature of our same-different task: When two items are
ambiguous but perceptually different, participants might have a slight
preference for \enquote{different} over \enquote{same}.

The values of the means are fixed: They correspond to the value of the
most typical tokens in our stimuli. Thus, observations from each
modality (\(a\) and \(v\)) are weighted according to their reliability
in the posterior: \[\beta_a \propto \frac{1}{\sigma^2_{A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}}\].

Note that the weights account for categorical variability only. For
instance, if the speaker generates a target production \(a_t\) from an
auditory category \(p(a_t | A) \sim N(\mu_{A}, \sigma^2_{A})\), the
ideal model assumes that it has direct access to this production token
(i.e., \(a=a_t\)), and that all uncertainty is about the category
membership of this token. However, we might also want to account for
noise in the brain and/or in the environment. For example, the observer
might not have access to the exact produced target, but only to the
target perturbed by noise. If we assume this noise to be normally
distributed, that is, \(p(a | a_t) \sim N(a_t, \sigma^2_{N})\), then
integrating over \(a_t\) lead to the following simple expression:
\[ p(a | A) \sim N(\mu_{A}, \sigma^2_{A}+\sigma^2_{N})\] Using an
identical reasoning for the case of the visual modality, we end up with
the following multimodal weighting scheme in the optimal combination
model:

\[\beta_a \propto \frac{1}{\sigma^2_{A}+\sigma^2_{N_A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V} +\sigma^2_{N_V}}\] To summarize,
the posterior provides the optimal model's predictions for how
probabilities that characterize uncertainty in each modality can be
combined to make categorical decision about the bimodal input. The model
will be constructed based on data from the unimodal trials, but the
predictions of the optimal model will be compared to human responses in
the bimodal trials.

\subsection{Auditory and Visual
models}\label{auditory-and-visual-models}

Besides the optimal model, we also test the predictions of two baselines
models: a visual model, which assumes that participants rely only on
visual information, and an auditory model, which assumes that
participants rely only on auditory information. If participants rely on
both the auditory and the visual modalities, the optimal model would
explain more variance in human responses than the visual model alone or
the auditory model alone.

\subsection{Descriptive model}\label{descriptive-model}

The visual, auditory, and optimal models are \emph{normative} models.
Their predictions are made about human data in the bimodal condition,
but their crucial parameters (i.e., variances associated with the visual
and auditory modalities) are derived from data in the unimodal
conditions. In addition to these normative models, we consider a
\emph{descriptive} model. The parameters of this model are fit to human
data in the bimodal condition. If the referential task induces
sub-optimality (due, for instance, to the arbitrary nature of the
sound-object association), then we predict that the descriptive model
should explain more variance than the optimal model does.

>>>>

\documentclass[english,floatsintext,man]{apa6}

@@

@@

How does the listener deal with uncertainty to identify the speaker's
intended word? One rigorous way to approach this question is through
conducting an \emph{ideal observer} analysis. This research strategy
provides a characterization of the task/goal and shows what the optimal
performance should be under this characterization.\footnote{It is, thus,
  a general instance of the rational approach to cognition (Anderson,
  1990). It can also be seens as an instance Marr's computational level
  of analysis.} When there is uncertainty in the input, the ideal
observer performs an optimal probabilistic inference. For example, in
order to recognize an ambiguous linguistic input, the model uses all
available probabilistic knowledge in order to maximize the accuracy of
this recognition. When the task is well specified, the ideal observer
model can be seen as a theoretical upper limit on performance. Thus, it
is supposed to be used, not so much as a realistic model of human
performance, as much as a baseline against which human performance can
be compared (Geisler, 2003; Rahnev \& Denison, 2018). When there is a
deviation from the ideal, it can reveal extra constraints on human
cognition, such as limitations on the working memory or attentional
resources.

The ideal observer analysis has had a tremendous impact not only on
speech related research (Clayards, Tanenhaus, Aslin, \& Jacobs, 2008;
Feldman, Griffiths, \& Morgan, 2009; Kleinschmidt \& Jaeger, 2015;
Norris \& McQueen, 2008), but also on many other disciplines in the
cognitive sciences (for reviews, see Chater \& Manning, 2006; Knill \&
Pouget, 2004; Tenenbaum, Kemp, Griffiths, \& Goodman, 2011). In
particular, using this research strategy, Clayards et al. (2008)
simulated auditory uncertainty by manipulating the probability
distribution of a cue (VOT) that differentiated similar words (e.g.,
\enquote{beach} and \enquote{peach}). They found that humans were
sensitive to these probabilistic cues and their judgments closely
reflected the optimal predictions. In another work, Feldman et al.
(2009) studied the perceptual magnet effect, which is a phenomenon that
involves reduced discriminability near prototypical sounds in the native
language (Kuhl, 1991). They showed that this effect can be explained as
the consequence of optimally solving the problem of perception under
uncertainty. Both Clayards et al. (2008) and Feldman et al. (2009)
explored optimal performance under uncertainty in the auditory modality.
There is, however, extensive evidence that information from the visual
modality, such as the speaker's facial features, also influences speech
understanding (see Campbell, 2008 for a review). Bejjanki, Clayards,
Knill, and Aslin (2011) offered a mathematical characterization of how
probabilistic cues from speech and lip movements can be optimally
combined. They showed that human performance during audio-visual
phonemic labeling was consistent (at least at the qualitative level)
with the behavior of the ideal observer.

This previous research, however, did not systematically study speech
understanding when the visual information is obtained, not through the
speaker's facial features, but through the referential context. In fact,
experimental findings showed that information about the identity of the
semantic referent can be integrated with linguistic information to
resolve lexical and syntactic ambiguities in speech (e.g., Eberhard,
Spivey-Knowlton, Sedivy, \& Tanenhaus, 1995; Spivey, Tanenhaus,
Eberhard, \& Sedivy, 2002; Tanenhaus, Spivey-Knowlton, Eberhard, \&
Sedivy, 1995). To our knowledge, however, no study offered an ideal
observer analysis of word identification in such context, that is, when
the listener has to combine cues from the sound and the referent.
Imagine, for example, that someone is uncertain whether they heard
\enquote{pea} or \enquote{bee}, does this uncertainty make them rely
more on the referent (e.g., the object being pointed at)? Vice versa, if
they are not sure if they saw a bee or a fly, does it make them rely
more on the sound? More importantly, when input in both modalities is
uncertain to varying degrees, do they weight each modality according to
its relative reliability (which is the optimal strategy), or do they
over-rely on a particular modality (which is a sub-optimal strategy)?

On the face of it, the question of combining information from the sound
and the visual referent might seem similar to that of audio-visual
speech integration. Nevertheless, there are at least two fundamental
differences between these two cases, and both can influence the way the
auditory and visual cues are integrated:

\noindent First, in the case of audio-visual speech, both modalities
offer information about the same underlying speech category. They may
differ only in terms of their informational reliability. In a
referential context, however, the auditory and visual modalities are
additionally different in terms of the roles they play in the
referential process: the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning}. It has been
suggested that because of its referential property, speech is a
privileged signal for humans, starting in infancy (see Vouloumanos \&
Waxman, 2014 for a review).\footnote{There is a debate as to whether
  speech is privileged for infant and adults for the same reasons.
  Whereas some researchers suggest that speech is privileged for both
  infants and adults because of its ability to refer (e.g., Waxman \&
  Markow, 1995), others suggest that speech might \emph{not} have a
  referential status from the start. Rather, speech might be prefered by
  infants only because of a low level auditory ``overshadowing'' (e.g.,
  Sloutsky \& Napolitano, 2003).} Thus, in a referential context, it is
possible that listeners do not treat the auditory and visual modalities
as equivalent sources of information. Instead, there could be a
sub-optimal bias for the auditory modality beyond what is expected from
informational reliability alone.

\noindent Second in the case of audio-visual speech, the auditory and
visual stimuli are expected to be perceptually correlated. The
expectation for this correlation is such that when there is a mismatch
between the auditory and visual input, people still integrate them into
a unified (but illusory) percept (McGurk \& MacDonald, 1976). In the
case of referential language, however, the multimodal association is by
nature \emph{arbitrary} (Greenberg, 1957; Saussure, 1916). For instance,
there is no logical/perceptual connection between the sound
\enquote{bee} and the corresponding insect. Moreover, variation in the
way the sound \enquote{bee} is pronounced is generally not expected to
correlate perceptually with variation in the shape (or any other visual
property) in the category of bees. In sum, cue combination in the case
of arbitrary audio-visual associations (word-referent) is likely to be
less automatic, more effortful, and therefore less conducive to optimal
integration than it is in the case of perceptually correlated
associations (as in audio-visual speech perception).

In the current study, we investigate how people combine cues from the
auditory and the visual modality to recognize words in a referential
context. In particular, we study how this combination is performed under
various degrees of uncertainty in both the auditory and the visual
modality. We perform a rational analysis of the task. First we propose
an ideal observer model that performs the combination in an optimal
fashion. Second we compare the predictions of the optimal model to human
responses. Humans can deviate from the ideal for several reasons. For
instance, as mentioned above, a sub-optimality can be induced by the
suggested privileged status of speech or by the arbitrariness of the
referential association. In order to study possible patterns of
sub-optimality, we compare the optimal model (which provides a normative
benchmark) to a descriptive model (which is fit to human responses).
Comparing parameter estimates between these two formulations allows us
to quantify the degree of deviation from optimality.

@@

\section{Paradigm and Models}\label{paradigm-and-models}

In this section we, first, briefly introduce the multimodal combination
task. Then we explain how behavior in this paradigm can be characterized
optimally with an ideal observer model.

@@

We introduce a new task that tests word recognition in a referential
context. We use two visual categories (cat and dog) and two auditory
categories (/b/ and /d/ embedded in the minimal pair /aba/-/ada/). For
each participant, an arbitrary pairing is set between the auditory and
the visual categories, leading to two audio-visual word categories
(e.g., dog-/aba/, cat-/ada/). In each trial, participants are presented
with an audio-visual target (the prototype of the target category),
immediately followed by an audio-visual test stimulus
(Figure~\ref{fig:task}). The test stimulus may differ from the target in
both the auditory and the visual components. After these two
presentations, participants press \enquote{same} or \enquote{different.}

@@

This paradigm is adapted from a previous task (Sloutsky \& Napolitano,
2003), which has been used with both children and adults to probe
audio-visual encoding (see Robinson \& Sloutsky, 2010 for a review). In
the testing phase of the original task, participants are asked whether
or not the two audio-visual presentations are \emph{identical}. In the
current study, we are interested, rather, in the categorization, i.e.,
determining whether or not two similar tokens are members of the same
phonological/semantic category. Therefore, testing in our task is
category-based: Participants are asked to press \enquote{same} if they
think the second item (the test) belongs to the same category as the
first (target) (e.g., dog-/aba/), even if there is a slight difference
in the sound, in the referent, or in both. They are instructed to press
\enquote{different} only if they think that the second stimulus was an
instance of the other category (cat-/ada/). The task also includes
trials where pictures are hidden (audio-only) or where sounds are muted
(visual-only). These unimodal trials provide us with the participants'
evaluation of the probabilistic information present in the auditory and
visual categories. As we shall see, these unimodal distributions are
used as inputs to the optimal cue combination model.

@@

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002)---which typically studies continuous stimuli (e.g., size,
location)---this probabilistic information cannot be characterized with
\emph{sensory noise/variability}, only. Indeed, our task involves
responses over categorical variables (phonemes and concepts), and
therefore, the optimal model in our case should take into account, not
only the noise variability around an individual perceptual estimate, but
also its \emph{within category variability}, i.e., the uncertainty
related to whether this perceptual estimate belongs to a given category
(see also Bankieris, Bejjanki, \& Aslin, 2017; Bejjanki et al., 2011).
In what follows, we describe a model that accounts for both type of
variability. First, we describe the model in the simplified case of
categorical variability only. Second, we augment this simplified model
to account for sensory noise.

\subsubsection{Categorical variability}\label{categorical-variability}

We assume that both the auditory categories (i.e., /aba/ and /ada/) and
the visual categories (cat and dog) are distributed along a single
acoustic and semantic dimension, respectively (Figure~\ref{fig:model}).
Moreover, we assume that all categories are normally distributed.
Formally speaking, if \(A\) denotes an auditory category (/ada/ or
/aba/), then the probability that a point \(a\) along the acoustic
dimension belongs to the category \(A\) is
\[ p(a | A) \sim  N(\mu_A, \sigma^2_A) \] where \(\mu_A\) and
\(\sigma^2_A\) are respectively the mean and the variance of the
auditory category. Similarly, the probability that a point \(v\) along
the visual dimension belongs to the category \(V\) is
\[ p(v | V) \sim  N(\mu_V, \sigma^2_V) \] where \(\mu_V\) and
\(\sigma^2_V\) are the mean and the variance of the visual category. An
audio-visual signal \(w=(a,v)\) can be represented as a point in the
audio-visual space. These audio-visual tokens define bivariate
distributions in the bi-dimentional space. We call these bivariate
distributions \emph{Word categories}, noted \(W\), and are distributed
as follows: \[ p(w | W) \sim  N(M_W, \Sigma_W) \] where
\(M_W=(\mu_A, \mu_V)\) and \(\Sigma_W\) are the mean and the covariance
matrix of the word category. The main assumption of the model is that
the auditory and visual variables are independent (i.e., uncorrelated),
so the covariance matrix is simply: \[
   \Sigma_W=
  \left[ {\begin{array}{cc}
   \sigma^2_A & 0 \\
   0 & \sigma^2_V \\
  \end{array} } \right]
\]

\begin{figure}[!h]
\includegraphics[width=\textwidth]{pictures/model} \caption{Illustration of the model using simulated data. A word category is defined as the joint bivariate distribution of an auditory category (horizontal, bottom panel) and a visual semantic category (vertical, left panel). Upon the presentation of a word token $w$, participants guess whether it is sampled from the word type $W_1$ or from the word type $W_2$. Decision threshold is where the guessing probability is 0.5.}\label{fig:model}
\end{figure}

\noindent This assumption simply says that, given a word-object mapping,
e.g., \(W=\)(\enquote{cat}-CAT), variation in the way \enquote{cat} is
pronounced does not correlate with changes in any visual property of the
object CAT, which is a valid
assumption.\footnote{Note that this assumptions is more adequate in the case of arbitrary associations such as ours, and less so in the case of redundant association such as audio-visual speech. In the latter, variation in the pronunciation is expected to correlate, at least to some extent, with lip movements.}

Now we turn to the crucial question of modeling how the optimal decision
should proceed given the probabilistic (categorical) information in the
auditory and the visual modalities, as characterized above. We have two
word categories: dog-/aba/ (\(W_1\)) and cat-/ada/
(\(W_2\)).\footnote{This mapping is randomized in the experiments.} When
making decisions, participants can be understood as choosing one of
these two word categories (Figure~\ref{fig:model}). For an ideal
observer, the probability of choosing category 2 when presented with an
audio-visual instance \(w=(a,v)\) is the posterior probability of this
category: \[
p(W_2 | w)=\frac{p(w|W_2)p(W_2)}{p(w|W_2)p(W_2)+p(w|W_1)p(W_1)}
\] Using our assumption that the cues are uncorrelated, we have:
\[p(w | W) = p(a,v| W) = p(a| A)p(v| V)\] Under this assumption, the
posterior probability reduces to the following formula (see Appendix 1
for the details of the derivation):

\begin{equation}
 p(W_2 | w)=\frac{1}{1+(1+b)\exp(\beta_0+\beta_aa+\beta_vv)}
\end{equation}

where \[1+b=\frac{p(W_1)}{p(W_2)}\]
\[\beta_0=\frac{\mu^2_{A2}-\mu^2_{A1}}{2\sigma^2_{A}}+\frac{\mu^2_{V2}-\mu^2_{V1}}{2\sigma^2_{V}}\]

\[\beta_a=\frac{\mu_{A1}-\mu_{A2}}{\sigma^2_{A}}\]
\[\beta_v=\frac{\mu_{V1}-\mu_{V2}}{\sigma^2_{V}}.\]

The parameter \(b\) represents the differential between the categories'
prior probabilities. However, since the identity of word categories is
randomized across participants, \(b\) measures, rather, a response bias
to \enquote{same} if \(b > 0\), and a response bias to
\enquote{different} if \(b < 0\). We expect a general bias towards
answering \enquote{different} because of the categorical nature of our
same-different task: When two items are ambiguous but perceptually
different, participants might have a slight preference for
\enquote{different} over \enquote{same}. As for the means, their values
are fixed, and they correspond to the most typical tokens in our
stimuli. Finally, observations from each modality (\(a\) and \(v\)) are
weighted in Equation 1 according to their reliability:
\[\beta_a \propto \frac{1}{\sigma^2_{A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V}}.\]

\subsubsection{Sensory variability}\label{sensory-variability}

So far, we only accounted for categorical variability. For instance, if
the speaker generates a target production \(a_t\) from an auditory
category \(p(a_t | A) \sim N(\mu_{A}, \sigma^2_{A})\), the ideal model
assumes that it has direct access to this production token (i.e.,
\(a=a_t\)), and that all uncertainty is about the category membership of
this token. However, we might also want to account for noise in the
brain and/or in the environment. For example, the observer might not
have access to the exact produced target, but only to the target
perturbed by noise. If we assume this noise to be normally distributed,
that is, \(p(a | a_t) \sim N(a_t, \sigma^2_{N_A})\), then integrating
over \(a_t\) leads to the following simple expression:
\[ p(a | A) \sim N(\mu_{A}, \sigma^2_{A}+\sigma^2_{N_A})\] Similarly, in
the case of sensory noise in the visual modality, we get
\[ p(a | V) \sim N(\mu_{V}, \sigma^2_{V}+\sigma^2_{N_V})\] Finally,
using exactly the same derivation as above, we end up with the following
multimodal weighting scheme in the optimal combination model (Equation
1) which takes into account both categorical and sensory variability:

\[\beta_a \propto \frac{1}{\sigma^2_{A}+\sigma^2_{N_A}}\]
\[\beta_v \propto \frac{1}{\sigma^2_{V} +\sigma^2_{N_V}}.\]

\subsubsection{Optimal cue combination}\label{optimal-cue-combination}

Equation 1 provides the optimal model's predictions for how
probabilities that characterize uncertainty in the auditory and the
visual modalities can be combined to make categorical decisions.
Parameters' estimates of the probability distributions in each modality
are derived by fitting unimodal posteriors to the participants'
responses in the unimodal conditions, i.e., the condition where only the
sounds are heard or only the pictures are seen
(Figure~\ref{fig:task}).\footnote{Further technical detail about model fitting in the unimodal conditions will be given in the method section of Experiment 1}
Using these derived parameters, the optimal model makes predictions
about responses in the bimodal condition where participants both hear
the sounds and see the pictures.

\subsubsection{Auditory and Visual
baselines}\label{auditory-and-visual-baselines}

The predictions of the optimal model will be compared to two baselines.
The first baseline is a visual model which assumes that participants
rely only on visual information, and an auditory model, which assumes
that participants rely only on auditory information. More precisely,
these baseline models assumes that the participants' responses in the
bimodal condition will not be different from their response in either
the visual-only or the auditory-only condition. If, as we expect, the
participants rely on both the auditory and the visual modalities to make
decision in the bimodal condition, the optimal model would explain more
variance in human responses than the visual or the auditory model do.

\subsection{Descriptive model and analysis of
sub-optimality}\label{descriptive-model-and-analysis-of-sub-optimality}

The optimal model (as well as the auditory and visual baselines) are
\emph{normative} models. Their predictions are made about human data in
the bimodal condition, but their crucial parameters (i.e., variances
associated with the visual and auditory modalities) are derived from
data in the unimodal conditions. In addition to these normative models,
we consider a \emph{descriptive} model. The parameters of this model are
fit to actual responses in the bimodal condition. If the referential
task induces sub-optimality (due, for instance, to the arbitrary nature
of the sound-object association), then the descriptive model should
explain more variance than the optimal model does.

Comparison of the optimal and the descriptive models allows us, not only
to quantify how much people deviate from optimality, but also to
understand precisely the nature of this deviation. Let \(\sigma^2_{A}\)
and \(\sigma^2_{V}\) be the values of the variances used in the optimal
model (derived from the unimodal conditions), and \(\sigma^2_{Ab}\) and
\(\sigma^2_{Vb}\) be the values observed through the descriptive model
in the bimodal condition. Deviation from optimality is measured in two
ways. First, we measure the change in the values of the variance
specific to each modality, that is, how \(\sigma^2_{A}\) compares to
\(\sigma^2_{Ab}\), and how \(\sigma^2_{V}\) compares to
\(\sigma^2_{Vb}\). Second, we measure changes in the proportion of the
visual and auditory variances, i.e., we examine how
\(\frac{\sigma^2_{A}}{\sigma^2_{V}}\) compares to
\(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}}\). The first measure allows us to
test if response precision changes for each modality when we move from
the unimodal to the bimodal conditions. The second allows us to test the
extent to which the weighting scheme follows the prediction of the
optimal model. The reason we used the proportion of the variances as a
measure of cross-modal weighting is because this proportion corresponds
to the slope\footnote{Or more precisely the absolute value of the slope}
of the decision threshold in the audio-visual space
(Figure~\ref{fig:model}). The decision threshold is defined as the set
of values in this audio-visual space along which the posterior is equal
to 0.5. Formally speaking, the decision threshold has the following
form:
